Scaling through abstractions – high-performance vectorial wave simulations for seismic inversion with Devito

### 2020-06-04 response.md

---
title: respons to reviews
bibliography:
	- sc20_paper.bib
---
Dear Editor,

First of all, we would like to thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.

@@

## R1

> run your model [...] How is that done?

During experimentation, we had two issues:
1 - Limited Azure resources, including access to an Infiniband-enabled system.
2 - Some MPI-level optimizations needed tuning and consolidation.

We managed to perform initial scaling experiments [@witte2019ecl;@luporini2019adp], although without the latest advancements in Devito (e.g., computation/communication overlap). Today, we have preliminary evidence that Devito-v4.2 has remarkable parallel efficiency, but publication would be premature.

Accuracy: details added to the text (in red). In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences. We also test through continuous integration the numerical accuracy of all implemented models.

@@

> [...] not very suitable for SC

This work was conceived as an application-oriented paper based on high-level abstractions and compilation technology. However, unlike many other works centered on DSLs, ours builds on -- and contributes to -- what is nowadays an established framework for seismic imaging, used in industry and academia. Our work shows the benefits (productivity, HPC, ...) of raising the level of abstraction, so it is relevant for SC.

> Parallelization is pretty straightforward [...]

Our work focuse on the generation of MPI-parallel code from a mathematical language for solving PDEs rather than the manual addition of domain decomposition toanexi  sting code. This requires considerable software engineering to allow users to run on a distributed-memory system with virtually no changes to their code. We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes unstructured sparse functions, implements distributed numpy arrays, and applies several optimazions for effective halo exchange.

> high arithmetic intensity [...] challenging for optimization

High arithmetic intensity does not imply "non-challenging optimization". High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization, notable yet misleasing GFlops/s performance is achieved since a huge fraction of flops are redundant (e.g., common sub-expressions). TTI is similar. As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtaine remarkable GFlops/s, but terrible GPoints/s. Optimizing away such redundancies is cumbersome, and is one of the reasons companies are so much interested in Devito.

@@

> [...] MPI which is a bit behind times these days.

MPI is still curently the most broadly accepted approach for domain decomposition.

@@

## R3

> little confusing but [...] style parallelism.

See R2 -- it's not just "grid decomposition", Devito deals with distribution of sparse functions, numpy arrays, decomposition over sub-domains, etc.

@@

> some mention of threads, [...]. Only MPI is detailed.

Correct. We explicitly refer to other articles for these other aspects.

> [...] FMA?

The operations determining the performance of a code are the source-level ones. FMA is a lower-level concept (ie, a scalar FMA = two operations). This model is adopted everywhere.

@@

> Scalability performance

See R1 -- (i) lack of resources and (ii) Devito not ready yet for scalability tests when the experimentation was performed.

> much of the performance [...]

We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations. Further, as explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system. To achieve this, we are using bleeding edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, the rapidly-spreading GitHub-Actions for CI, etc. The use and orchestration of these technologies is thouroughly described in the paper, and is of strong interest for SC (strong industry interest for production scale).

@@

> For actual seismic [...] it helps a little.

The industrial problem does indeed involve muliple parts, such as data acquisition, processing, and, most importantly, velocity model building. We do not claim to solve all these problems. However, we demonstrate that the combination of advanced technolgies such as DSLs and compilers is key to highly-efficient sesimic modelling and imaging at industrial scale. We have added references and clarification in the introduction.

> Dropping [...] 3D elastic [...] might demonstrate [...].

The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as important as a strong scaling experiment. Strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:

### 2020-06-05 response.md

---
title: response to reviews
bibliography:
	- sc20_paper.bib
---
Dear Editor,

We thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.

## R1
> If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that? 
We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp]. While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture. Applications such as FWI generally only use MPI for computing on a single data shot (i.e. a single observation). However, that computation unit of work sits within an embarrassingly parallel loop processing thousands of such shots as part of a gradient-decent algorithm. So while the computation of a single shot (the focus on Devito_ might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS. A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that do describe the underlying algorithm.

> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?
Details on solver verification have been added to the paper (in red). In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences. We also test through continuous integration the numerical accuracy of all implemented models. Additionally, for inverse problems, we use the standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian) [@louboutin2018segeow; @louboutin2020THmfi]"

@@

## R2

> [...] not very suitable for SC
There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable for SC. High productivity for HPC and performance-portability continue to be hot topics at SC. Many recent Gordon Bell finalists in recent years have used to some degree DSL's, and DSL's. Devito is one of the very examples of DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale. While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain because of the commercial sensitivity. However, one can conclude from the fact that Devito is being adopted by companies commercially to suggest comparisons were favourable. The research in Devito is highly relevant to SC as it demonstrates that the HPC cliche that one have to can have high-productivity languages or high-performance but not both is false.

@@

> Parallelization is pretty straightforward [...]

Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra... 
The novelty is that our work involves developing a layers of abstracts that allow finite difference based solvers be to implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism. This is not "straightforward" - the only comparable software technologies that we are aware of to have accomplished something similar are FEnICS and Firedrake for the finite-element method.
We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed numpy arrays, and applies several optimizations for efficient halo exchange.

@@

> high arithmetic intensity [...] challenging for optimization

High arithmetic intensity does not imply "non-challenging optimization". High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization results in an inflated GFlops/s measurement since a huge fraction of flops are redundant (e.g., common sub-expressions). TTI, a relatively sophisticated propagation model used in industry, is similar. As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution). Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand as there can be thousands of algebraic terms in these engineering applications.

@@

> [...] MPI which is a bit behind times these days.

The authors are curious to know if the reviewer also believes all MPI focused papers are unsuitable for SC. Devito uses a range of parallel programming models, usually nested, depending on the context. For example, SIMD for SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs. For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.

@@

## R3

> little confusing but [...] style parallelism.

The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns. While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation, distributed numpy arrays.

@@

> some mention of threads, [...]. Only MPI is detailed.

Correct. We explicitly refer to other articles for these other aspects.

> [...] FMA?

The operations determining the performance of software are the source-level ones. FMA is a lower-level concept (i.e., a scalar FMA = two operations). This model is adopted everywhere.

@@

> Scalability performance

See response to R1.

> much of the performance [...]

We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations. Further, as explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system. To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system such as Devito, Azure cloud, Docker, GitHub-Actions for CI, etc. The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest for SC (strong industry interest for production scale).

@@

> For actual seismic [...] it helps a little.

The industrial problem does indeed involve multiple parts, such as data acquisition, processing, and, most importantly, velocity model building. We do not claim to solve all these problems. However, we demonstrate that the combination of advanced technologies such as DSLs and compilers is key to highly-efficient seismic modelling and imaging at an industrial scale. We have added references and clarification in the introduction.

@@

> Dropping [...] 3D elastic [...] might demonstrate [...].

The new capabilities of Devito -- tensor language, MPI from high-level abstractions -- for multiple physics and discretizations is as essential as a strong scaling experiment. Strong scaling may, to some extent, be regarded as more of an academic exercise in seismic imaging:

@@

> section III describing new work?

Yes. We have improved the text to make this clear.

> parallelism just via MPI?

The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition). In this context, task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II]. This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.

@@

> What are the possibilities [...]

GPU support is in development and currently (Devito-v4.2) supported for some applications. Its capabilities and related applications will be covered in a forthcoming paper.

### 2020-06-05 response.md

---
title: response to reviews
bibliography:
	- sc20_paper.bib
---
Dear Editor,

We thank the reviewers for their work and constructive feedback. Below, we address their questions and observations. All edits are in red in the manuscript.

## R1
> If you run your model with a variable number of cores, how is the weak and strong scaling? Can you comment on that?

We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp]. While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture. Applications such as FWI generally only use MPI for computing on a single data shot (i.e. a single observation). However, that computation unit of work sits within an embarrassingly parallel loop processing thousands of such shots as part of a gradient-decent algorithm. So while the computation of a single shot (the focus on Devito might only require TFLOPS, and iteration of the outer loop would be typically PFLOPS). A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that do describe the underlying algorithm.

> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality? How is that done?

Details on solver verification have been added to the paper (in red). In [@devito-api], we analyze the numerical accuracy against the analytical solution and the convergence rate of the finite-differences. We also test through continuous integration the numerical accuracy of all implemented models. Additionally, for inverse problems, we use the standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian) [@louboutin2018segeow; @louboutin2020THmfi]"

@@

## R2

> [...] not very suitable for SC

There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable for SC. High productivity for HPC and performance-portability continue to be hot topics at SC. Many Gordon Bell finalists in recent years have used to some degree DSL's. Devito is one of the very examples of DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale. While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain because of the commercial sensitivity. However, one can conclude from the fact that Devito is being adopted by companies commercially to suggest comparisons were favourable. The research in Devito is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.

@@

> Parallelization is pretty straightforward [...]

Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra... 
The novelty is that our work involves developing layers of abstraction that allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism. This is not "straightforward" - the only comparable software technologies that we are aware of to have accomplished something similar are FEnICS and Firedrake for the finite-element method.
We also point out in the paper [Sections I and II-B], that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.

@@

> high arithmetic intensity [...] challenging for optimization

High arithmetic intensity does not imply "not as challenging for optimization". High-order finite (spectral) elements for instance -- where an extremely high arithmetic intensity can be obtained, but without tensor-product basis functions and optimizations such as sum-factorization results in an inflated GFlops/s measurement since a huge fraction of flops are redundant (e.g., common sub-expressions). TTI, a relatively sophisticated propagation model used in industry, is similar. As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution). Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand as there can be thousands of algebraic terms in these engineering applications.

@@

The authors are curious to know if the reviewer also believes all MPI focused papers are unsuitable for SC. Devito uses a range of parallel programming models, usually nested, depending on the context. For example, SIMD for SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs. For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.

## R3

> little confusing but [...] style parallelism.

The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns. While we focus on MPI  for domain decomposition for this paper, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.

@@

> parallelism just via MPI?

The inversion (outer-level of parallelism) requires what is essentially task parallelism; each task is internally MPI-parallel (domain decomposition). In this context, task-parallelism means parallelization over sources, accomplished using `batch-shipyard` in a serverless setting [Section II]. This allowed optimal usage of the cloud, where one pays based on actual usage of computational resources.

> What are the possibilities [...]

GPU support is in development and currently (Devito-v4.2) covers some applications. Its capabilities and related applications will be covered in a forthcoming paper.

### 2020-06-05 response.md

---
title: response to reviews
bibliography:
	- sc20_paper.bib
---
Dear Editor,

We thank the reviewers for their work and constructive feedback. Below we address their questions and observations. Edits marked in red.

## R1
> If you run your model with a variable number of cores, how is the weak and strong scaling?

We published and cited MPI scaling results in [@witte2019ecl;@luporini2019adp]. While Devito MPI scaling has improved significantly since those publications, it is only a relatively minor consideration within the bigger picture. Applications such as FWI generally only use MPI for computing a shot (single observation). That unit of work sits within an embarrassingly parallel loop processing thousands of shots as part of a gradient-descent algorithm. While computing a single shot the focus on Devito might only require TFLOPS -- iteration of the outer loop would typically be PFLOPS. A description of nested-parallelism in applications such as FWI is outside of the scope of this paper, but we have included several references to FWI papers that describe the underlying algorithm.

> Further, you mention that you validate performance against a wave propagator - how about validation of simulation quality?

Details on solver verification have been added to the paper. In [@devito-api], we analyze the numerical accuracy against analytical solutions and the convergence rate of the finite-differences. We also test through continuous integration the numerical accuracy of all implemented models. For inverse problems, we use standard dot and gradient tests, "we focused on correct and testable implementations for the adjoint wave-equation and the gradient (action of the adjoint Jacobian)" [@louboutin2018segeow; @louboutin2020THmfi].

@@

## R2

> [...] not very suitable for SC

There are few words in this review so we can only guess in what sense the reviewer feels the work is not suitable. High productivity for HPC and performance-portability continue to be hot topics at SC. Many Gordon Bell finalists in recent years have used to some degree DSL's. Devito is one of the few DSL's to have challenged the performance of handwritten HPC codes for real engineering applications at scale. While Devito is fully open source, we have not been able to publish any head-to-head comparisons with commercial codes in this domain owing to commercial sensitivity. However, one can conclude from the fact that Devito is being adopted by companies commercially that comparisons were favourable. The research is highly relevant to SC as it demonstrates the HPC cliche that one can have high-productivity languages or high-performance, but not both, is false.

@@

> Parallelization is pretty straightforward [...]

Whether or not parallelization is straightforward is not the issue. One can make the same comment about any finite difference, stencil code, anything involving linear algebra...
The novelty our work is that it involves developing layers of abstraction to allow finite difference based solvers to be implemented in symbolic mathematical form and automatically generate optimized C complete with nested MPI, OpenMP and SIMD parallelism. This is not "straightforward" - the only comparable software technologies that we are aware of are FEnICS and Firedrake.
We also point out in the paper [Sections I and II-B] that on top of standard domain decomposition, Devito decomposes irregular and unstructured sparse functions, implements distributed NumPy arrays, and applies several optimizations for efficient halo exchange.

@@

> high arithmetic intensity [...] challenging for optimization

High arithmetic intensity does not imply "not as challenging for optimization". Take high-order finite elements for instance -- extremely high arithmetic intensity can be easily obtained, but without tensor-product basis functions and optimizations such as sum-factorization results (GFlops/s) are inflated since a huge fraction of flops are redundant (e.g., common sub-expressions). TTI, a relatively sophisticated propagation model used in industry, is similar. As explained in Section III, without capturing cross-iteration common sub-expressions (high-order cross derivatives), we would obtain very high GFlops/s, but terrible GPoints/s (and therefore poor time to solution). Optimizing away such redundancies, one of the compiler transformations performed by Devito,  is challenging to do by hand since there can be thousands of algebraic terms.

@@

> [...] MPI which is a bit behind times these days.

The authors are curious to know if the reviewer also believes all MPI focused papers are unsuitable for SC. Devito uses a range of parallel programming models, usually nested, depending on the context. For example, SIMD registers, OpenMP for multi-core shared memory, OpenMP 5 offloading and OpenACC for accelerators, and frameworks such as Dask for parallelizing task-graphs. For domain-decomposition parallelism on distributed memory computers, MPI is still the most widely accepted solution.

@@

## R3

> little confusing but [...] style parallelism.

The complexity arises from the fact that there are several layers of parallelism and parallel programming patterns. While we here focus on MPI for domain decomposition, a single application will also use threading, accelerators, SIMD vectorization, irregular computation and distributed numpy arrays.

@@

> some mention of threads, [...]. Only MPI is detailed.

Correct. We explicitly refer to other articles for other aspects.

> [...] FMA?

Operations determining the performance of software are the source-level operations. FMA is a lower-level concept (scalar FMA = two operations). This model is widely adopted.

@@

> Scalability performance

See R1 response.

> much of the performance [...]

We show performance results of three different codes -- TTI, elastic, and isotropic acoustic -- thus exploring different physics and discretizations. As explained in R2, ours is an application-oriented paper built off high-level abstractions running on a cloud-based system. To achieve this, we are using state-of-the-art edge technologies -- an established DSL-based system, Azure cloud, Docker, GitHub-Actions for CI, etc. The use and orchestration of these technologies is thoroughly described in the paper and is of keen interest to SC (strong industry interest for production scale).

@@

> For actual seismic [...] it helps a little.

The industrial problem does indeed involve multiple parts, including data acquisition, processing, and, most importantly, velocity model building. We do not claim to solve all these problems. However, we demonstrate that the combination of advanced technologies such as DSLs and compilers is key to highly-efficient seismic modelling and imaging at industrial scale. We have added references and clarification in the introduction.

@@

* even with high-frequency FWI, a typical shot-level computation requires no more than a bunch of nodes (often order of units);
* there is an outer-level of parallelism for inversion, essentially a task farm where the individual workers are internally MPI-parallel.

## R4

> [...] how much of this work is new

We have improved the manuscript to clarify new contributions. These are also reiterated throughout the whole paper.

@@

> [...] vector and tensors [...] unclear [...] involved.

Short answer is significant: https://github.com/devitocodes/devito/pull/873

### 2020-06-05 paper.tex
rhodrin final tweaks (typos + grammar).

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem sizes
as encountered in exploration geophysics. These experiments are carried
out in Cloud-based HPC systems and include elastic forward modelling using
distributed-memory parallelism and imaging based on the tilted transverse
isotropic (TTI) wave-equation (\cite{virieux, thomsen1986, zhang-tti,
duveneck, louboutin2018segeow}). These proof of concepts highlight two
critical features: first, the ability of the symbolic interface and the
compiler to translate to large-scale adjoint-based inversion problems that
require massive compute (since thousands of PDEs are solved) as well as
large amounts of memory (since the adjoint state method requires the
forward model to be saved in memory). Secondly, through the elastic
modelling example, we demonstrate that \devito now fully supports and
automates vectorial and second order tensorial staggered-grid
finite-differences with the same high-level API previously presented for
scalar fields defined on cartesian grids.

This paper is organized as follows: first, we provide a brief overview of
Devito and its symbolic API and present the \textcolor{red}{newly developped} distributed memory
implementation that allows large-scale modelling and inversion by means of
domain decomposition. We then provide a brief comparison with a state of
the art hand-coded wave propagator to validate the performance previously
benchmarked with the roofline model (\cite{patterson, devito-compiler,
devito-api, louboutin2016ppf}). \textcolor{red}{Before concluding, results from the
Cloud-based experiments introduced above are presented}.

@@

Devito \cite{devito-api} provides a functional language built on top of
\texttt{SymPy} \cite{sympy} to symbolically express PDEs at a
mathematical level and implements automatic discretization with the
finite-difference method. The language is by design flexible enough to
express other types of non finite-difference operators, such as
interpolation and tensor contractions, that are inherent to
measurement-based inverse problems. Several additional features are
supported, among which are staggered grids, sub-domains, and stencils with
custom coefficients. The last major building block of a solid PDE solver
are the boundary conditions which for finite-difference methods are
notoriously diverse and often complicated. The system is, however,
sufficiently general to express them through a composition of core
mechanisms. For example, free surface and perfectly-matched \textcolor{red}{layer
(PML) conditions} can be expressed as equations -- just like
any other PDE equations -- over a suitable sub-domain.

@@

In this section we illustrate the \devito language by demonstrating
an implementation of the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(xs) q(t) \\
 u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
 d(t, .) = u(t, xr).
 \end{cases}
\label{acou}
\end{equation}
%
\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $xs$ is the point source position, 
$q(t)$ is the source time signature, $d(t, .)$ is the measured data at positions $xr$ and $m(x)$ is the
squared slowness.}
 The core of the Devito
symbolic API consists of three classes:

@@

\subsection{3D Imaging example on Azure}\label{d-imaging-example-on-azure}

@@

While the subsurface image obtained in section \ref{d-imaging-example-on-azure}
utilized anisotropic propagators capable of mimicking intricate physics, in order
to model both the wave kinematics and amplitudes correctly, elastic
propagators are required. These propagators are, for example, extremely
important in global seismology since shear surface waves (which are
ignored in acoustic models) are the most hazardous. \textcolor{red}{In this section, we 
introduce the new tensor algebra language implemented in \devito that enables the simple and 
high-level definition of vectorial and tensorial physics. This API extension also implements 
fully automatized staggered grid and staggered ginite-difference generally required for a stable discretization
of vectorial PDEs. This extension allows us to express an elastic model with compact and elegant notation 
and model its solution computationally efficiently with the \devito compiler}

@@

Transitioning from academic toy problems, such as the two-dimensional
acoustic wave-equation, to real-world applications can be challenging,
particularly if this transition is carried out as an afterthought.
Owing to the fundamental design principles of \devito such scaling,
however, becomes trivial. In this work we demonstrated
the high-level interface provided by \devito not only for simple scalar
equations but also for coupled PDEs. This interface allows, in a simple, concise
and consistent manner, the expression of all-kinds of non-trivial
differential operators. Next, and most
importantly, \textcolor{red}{it was} demonstrated that the compiler enables large-scale modelling with
state-of-the art computational performance and programming paradigm. The
single-node performance is on par with state of the art hand-coded models, but
packaged with this performance comes
the flexibility of the symbolic interface and multi-node parallelism, which is
integrated in the compiler and interface in a accessible way. Finally,
we demonstrated that our abstractions provide the necessary portability
to enable both on-premise and Cloud based HPC.

@@

\begin{itemize}
  \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019/tree/v1.0} and includes 
  the TTI propagators, the Azure setup for shot parallelism and a documentation.
  \item The elastic modelling can be run with the elastic example available in Devito at
  \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py} 
  and can be run with any size and spatial order. A standalone script to run the large 3D elastic modelling
  is also available at \url{https://github.com/mloubout/SC20Paper/tree/master/codesamples}
\end{itemize}

>>>>

Previous work focused on the DSL and compiler to highlight the potential
application and use cases of Devito. Here, we present a series of
extensions and applications to large-scale three-dimensional problem sizes
as encountered in exploration geophysics. These experiments are carried
out in Cloud-based HPC systems and include elastic forward modelling using
distributed-memory parallelism and imaging based on the tilted transverse
isotropic (TTI) wave-equation~\cite{virieux, thomsen1986, zhang-tti,
duveneck, louboutin2018segeow}. These proof of concepts highlight two
critical features: first, the ability of the symbolic interface and the
compiler to translate to large-scale adjoint-based inversion problems that
require massive compute (since thousands of PDEs are solved) as well as
large amounts of memory (since the adjoint state method requires the
forward model to be saved in memory). Secondly, through the elastic
modelling example, we demonstrate that \devito now fully supports and
automates vectorial and second order tensorial staggered-grid
finite-differences with the same high-level API previously presented for
scalar fields defined on cartesian grids.

This paper is organized as follows: first, we provide a brief overview of
Devito and its symbolic API and present the \textcolor{red}{newly developped} distributed memory
implementation that allows large-scale modelling and inversion by means of
domain decomposition. We then provide a brief comparison with a state of
the art hand-coded wave propagator to validate the performance previously
benchmarked with the roofline model \cite{patterson, devito-compiler,
devito-api, louboutin2016ppf}. \textcolor{red}{Before concluding, results from the
Cloud-based experiments introduced above are presented}.

@@

Devito \cite{devito-api} provides a functional language built on top of
\texttt{SymPy} \cite{sympy} to symbolically express PDEs at a
mathematical level and implements automatic discretization with the
finite-difference method. \textcolor{red}{The language is by design flexible enough to
express other non finite-difference operators, such as
interpolation and tensor contractions, inherent in
measurement-based inverse problems}. Several additional features are
supported, among which are staggered grids, sub-domains, and stencils with
custom coefficients. The last major building block of a solid PDE solver
are the boundary conditions which for finite-difference methods are
notoriously diverse and often complicated. The \textcolor{red}{DSL} is, however,
sufficiently general to express them through a composition of core
mechanisms. For example, free surface and perfectly-matched \textcolor{red}{layer
(PML) conditions} can be expressed as equations -- just like
any other PDE equations -- over a suitable sub-domain.

@@

In this section we illustrate the \devito language by demonstrating
an implementation of the acoustic wave-equation in isotropic media
%
\begin{equation}
\begin{cases}
 m(x) \frac{\partial^2 u(t, x)}{\partial t^2} - \Delta u(t, x) = \delta(x_s) q(t) \\
 u(0, .) = \frac{\partial u(t, x)}{\partial t}(0, .) = 0 \\
 d(t, .) = u(t, x_r).
 \end{cases}
\label{acou}
\end{equation}
%
\textcolor{red}{where $u(t, x)$ is the pressure wavefield, $x_s$ is the point source position, 
$q(t)$ is the source time signature, $d(t, .)$ is the measured data at positions $x_r$ and $m(x)$ is the
squared slowness.}
 The core of the Devito
symbolic API consists of three classes:

@@

\subsection{3D Imaging on Azure}\label{d-imaging-example-on-azure}

@@

While the subsurface image obtained in section \ref{d-imaging-example-on-azure}
utilized anisotropic propagators capable of mimicking intricate physics, in order
to model both the wave kinematics and amplitudes correctly, elastic
propagators are required. These propagators are, for example, extremely
important in global seismology since shear surface waves (which are
ignored in acoustic models) are the most hazardous. \textcolor{red}{In this section, we 
introduce the new tensor algebra language implemented in \devito that enables the simple and 
high-level definition of vectorial and tensorial physics. This API extension also implements 
fully automatized staggered grid and staggered finite-differences which are generally required for a stable discretization
of vectorial PDEs. This extension allows us to express an elastic model with compact and elegant notation 
and model its solution computationally efficiently with the \devito compiler}.

@@

Transitioning from academic toy problems, such as the two-dimensional
acoustic wave-equation, to real-world applications can be challenging,
particularly if this transition is carried out as an afterthought.
Owing to the fundamental design principles of \devito such scaling,
however, becomes trivial. In this work we demonstrated
the high-level interface provided by \devito not only for simple scalar
equations but also for coupled PDEs. This interface allows, in a simple, concise
and consistent manner, the expression of all-kinds of non-trivial differential operators\textcolor{red}{: the vectorial elastic equation 
and TTI (rotated finite-difference) being two examples presented in the present study}. Next, and most
importantly, \textcolor{red}{it was} demonstrated that the compiler enables large-scale modelling with
state-of-the art computational performance and programming paradigm. The
single-node performance is on par with state of the art hand-coded models, but
packaged with this performance comes
the flexibility of the symbolic interface and multi-node parallelism, which is
integrated in the compiler and interface in a accessible way. Finally,
we demonstrated that our abstractions provide the necessary portability
to enable both on-premise and Cloud based HPC.

\textcolor{red}{We emphasize that the work
presented here does not solve (and does not claim to solve) all problems encountered within exploration geophysics. The high-level interface 
and at (industry) scale performance demonstrated has resulted in
multiple academic and industrial projects adopting the use \devito within their larger imaging frameworks.}

@@

\begin{itemize}
  \item The complete code for TTI imaging is available at \url{https://github.com/slimgroup/Azure2019/tree/v1.0} and includes 
  the TTI propagators, the Azure setup for shot parallelism and documentation.
  \item \textcolor{red}{Elastic simulations can be run using the elastic example} available in Devito at
  \url{https://github.com/devitocodes/devito/blob/v4.2/examples/seismic/elastic/elastic_example.py}. A standalone script to run the large 3D elastic modelling \textcolor{red}{experiment presented here}
  is also available at \url{https://github.com/mloubout/SC20Paper/tree/master/codesamples}
\end{itemize}
